---
title: "R Assignment 3"
author: "Keeno Glanville"
date: "2022-07-28"
output: html_document
---

```{r setup, include=FALSE}
library(compare)
library(readr)
library(tidyverse)
library(dplyr)
library(corrplot)
library(car)
library(caret)
library(lmtest)
library(lattice)
library(PerformanceAnalytics)
```

# What factor influences crime?
### Load data(bonus)

```{r}
set.seed(1324)
urldata <- "https://raw.githubusercontent.com/vincentarelbundock/Rdatasets/master/csv/AER/Guns.csv"
gundata.r <- read_csv(url(urldata))



```



## We first start by cleaning the data set and renaming columns to make the names clearer
```{r}
gundata_clean <- gundata.r %>%
  select(-"...1", -"male")%>%
  rename("%Black" = "afam",
         "%White"= "cauc",
         "Crime/100k" = "violent",
         "Murder/100k" = "murder",
         "Robbery/100k" = "robbery",
         "Prisoners/100k" = "prisoners",
         "Income/capitaofstate" = "income",
         "Density(sqrmi/1000)" = "density",
          ShallCarryLaw = "law",
         "population/mil" = population)
view(gundata_clean)



```


### We wil now pair the data together to recognize the trends within our data

```{r}
gundata_pairs <- gundata_clean%>%
  mutate(ShallCarryLaw = ifelse(ShallCarryLaw == "no",0,1))
pairs(data=gundata_pairs,
      ~ year+ `Crime/100k`+ `Murder/100k` + `Robbery/100k`+ `Prisoners/100k`+
        `Density(sqrmi/1000)` + `Income/capitaofstate`)

```

## In the before step we see tremendous trends within the data and now we want to see our correlation coefficients
### We will print corrplot and histograms as well as boxplots

```{r}
gundata_corplot <- gundata_clean %>%
  select(-"state", -"ShallCarryLaw")

chart.Correlation(gundata_corplot,
                  method="pearson",
                  histogram=TRUE,
                  pch=16)

gundata_cor <- cor(gundata_corplot, method = "pearson")
col_gd <- colorRampPalette(c("#BB4444", "#EE9988", "#FFFFFF", "#77AADD", "#4477AA"))
corrplot(gundata_cor, method = "color", col = col_gd(200),
         type = "upper", order = "hclust",
         addCoef.col = "black",
         tl.col = "black", tl.srt = 45,)

```



# BoxPlot shows the trend of crime over the years, with a slight trend upwards heavily skewed by District of Columbia

```{r}

ggplot(gundata_clean, aes(x= year, y= `Crime/100k`, group = `year`)) +
         geom_boxplot( fill="skyblue", notch=FALSE) +
         geom_jitter( size=1, color="orange", width=0.3)

```




### Scatter Density vs Murder as this has our largest correlation coefficient 

```{r}
attach(gundata_clean)
DensityvMurder <- lm(`Murder/100k` ~ `Density(sqrmi/1000)`, data = gundata_clean)
plot(DensityvMurder, ask =FALSE)
avPlots(DensityvMurder)
detach(gundata_clean)

```



## Now we want to look at plots of crime with regards to state. We see District of Columbia elevated overall
```{r}
xyplot(log(`Crime/100k`) ~ as.numeric(as.character(year)) | state, data = gundata_clean, type = "l")

```

# We conclude that we see the data is heavily skewed due to the District of Columbia
```{r}
summary(DensityvMurder)

```


# We want to compare the data again but removing District of Columbia as a huge outlier




### Plots redone without District of Columbia
```{r}
attach(gundata_clean)
gundata.noDOC <- filter(gundata_clean, state != "District of Columbia")
gundata_noDOC <- gundata.noDOC %>%
  select(-"state", -"ShallCarryLaw")
detach(gundata_clean)

pairs(data=gundata_noDOC,
      ~ year+ `Crime/100k`+ `Murder/100k` + `Robbery/100k`+ `Prisoners/100k`+
        `Density(sqrmi/1000)` + `Income/capitaofstate`)

chart.Correlation(gundata_noDOC,
                  method="pearson",
                  histogram=TRUE,
                  pch=16)

gundata_noDOC_cor <- cor(gundata_noDOC,  method = "pearson")
col_noDOC <- colorRampPalette(c("#BB4444", "#EE9988", "#FFFFFF", "#77AADD", "#4477AA"))
corrplot(gundata_noDOC_cor, method = "color", col = col_noDOC(200),
         type = "upper", order = "hclust",
         addCoef.col = "black",
         tl.col = "black", tl.srt = 45,)


ggplot(gundata_noDOC, aes(x= year, y= `Crime/100k`, group = `year`)) +
  geom_boxplot( fill="skyblue", notch=FALSE) +
  geom_jitter( size=1, color="orange", width=0.3)



```







# We still see stronger parameters and correlation even with District of Columbia removed as an outlier

```{r}
attach(gundata_noDOC)
DensityvMurder_noDOC <- lm(`Murder/100k` ~ `Density(sqrmi/1000)`, data = gundata_noDOC)
plot(DensityvMurder_noDOC, ask =FALSE)
avPlots(DensityvMurder_noDOC)
detach(gundata_noDOC)


```




# From the data we see that no one factor influences Crime rates within a given city, it more so stems from a factor of all parameters






